\documentclass[11pt,twoside]{article}

\usepackage[T1]{fontenc}
\usepackage{newpxtext}
\usepackage{amsmath,amsthm,mathtools}
\usepackage{newpxmath}
\usepackage{booktabs}
\usepackage[
  activate={true,nocompatibility},
  tracking=true,
  expansion=true,
  factor=1100,
  stretch=10,
  shrink=10
]{microtype}
\microtypecontext{spacing=nonfrench}
\usepackage{geometry}
\usepackage[dvipsnames]{xcolor}
\usepackage{fancyhdr}
\usepackage{titlesec}
\usepackage{setspace}
\usepackage{needspace}
\usepackage{etoolbox}
\usepackage[most]{tcolorbox}

\geometry{
  paperwidth=8.5in,
  paperheight=11in,
  left=1.2in,
  right=1.2in,
  top=1.2in,
  bottom=1.2in,
  headheight=15pt
}

\widowpenalty=10000
\clubpenalty=10000
\raggedbottom

\setlength{\abovedisplayskip}{12pt plus 3pt minus 4pt}
\setlength{\belowdisplayskip}{10pt plus 3pt minus 4pt}
\setlength{\abovedisplayshortskip}{6pt plus 3pt}
\setlength{\belowdisplayshortskip}{8pt plus 3pt minus 3pt}

\definecolor{sectioncolor}{RGB}{25,25,45}
\definecolor{linkcolor}{RGB}{60,50,130}
\definecolor{citecolor}{RGB}{100,50,50}
\definecolor{theorembar}{RGB}{50,50,80}
\definecolor{defnbg}{RGB}{245,243,248}
\definecolor{remarkbar}{RGB}{140,140,160}

\titleformat{\section}
  {\normalfont\Large\bfseries\color{sectioncolor}}
  {\thesection}{1em}{}
\titleformat{\subsection}
  {\normalfont\large\bfseries\color{sectioncolor}}
  {\thesubsection}{1em}{}
\titleformat{\subsubsection}
  {\normalfont\normalsize\bfseries\color{sectioncolor}}
  {\thesubsubsection}{1em}{}

\titlespacing*{\section}{0pt}{3.5ex plus 1.5ex minus 1ex}{2ex plus .2ex}
\titlespacing*{\subsection}{0pt}{2.5ex plus 1ex minus .5ex}{1.2ex plus .2ex}
\titlespacing*{\subsubsection}{0pt}{2ex plus 0.8ex minus .3ex}{1ex plus .2ex}

\pretocmd{\section}{\needspace{6\baselineskip}}{}{}
\pretocmd{\subsection}{\needspace{5\baselineskip}}{}{}

\pagestyle{fancy}
\fancyhf{}
\fancyhead[LE,RO]{\small\thepage}
\fancyhead[RE]{\small\itshape Microprice and Semi-Markov Reinforcement Learning}
\fancyhead[LO]{\small\itshape L.\,Petersen}
\renewcommand{\headrulewidth}{0.3pt}
\renewcommand{\footrulewidth}{0pt}

\fancypagestyle{plain}{
  \fancyhf{}
  \fancyfoot[C]{\small\thepage}
  \renewcommand{\headrulewidth}{0pt}
}

\theoremstyle{plain}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{corollary}[theorem]{Corollary}

\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}

\theoremstyle{remark}
\newtheorem{remark}[theorem]{Remark}

\tcolorboxenvironment{theorem}{
  blanker, borderline west={2.5pt}{0pt}{theorembar},
  left=10pt, before skip=10pt plus 2pt minus 2pt,
  after skip=10pt plus 2pt minus 2pt, breakable
}
\tcolorboxenvironment{proposition}{
  blanker, borderline west={2.5pt}{0pt}{theorembar},
  left=10pt, before skip=10pt plus 2pt minus 2pt,
  after skip=10pt plus 2pt minus 2pt, breakable
}
\tcolorboxenvironment{lemma}{
  blanker, borderline west={2.5pt}{0pt}{theorembar},
  left=10pt, before skip=10pt plus 2pt minus 2pt,
  after skip=10pt plus 2pt minus 2pt, breakable
}
\tcolorboxenvironment{corollary}{
  blanker, borderline west={2.5pt}{0pt}{theorembar},
  left=10pt, before skip=10pt plus 2pt minus 2pt,
  after skip=10pt plus 2pt minus 2pt, breakable
}
\tcolorboxenvironment{definition}{
  colback=defnbg, colframe=defnbg, boxrule=0pt, arc=2pt,
  left=8pt, right=8pt, top=6pt, bottom=6pt,
  before skip=10pt plus 2pt minus 2pt,
  after skip=10pt plus 2pt minus 2pt, breakable
}
\tcolorboxenvironment{remark}{
  blanker, borderline west={1.5pt}{0pt}{remarkbar},
  left=10pt, before skip=8pt plus 2pt minus 2pt,
  after skip=8pt plus 2pt minus 2pt, breakable
}

\AtEndEnvironment{proof}{\medskip}

\newcommand{\1}{\mathbf{1}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\E}{\mathbb{E}}
\newcommand{\Prob}{\mathbb{P}}
\newcommand{\norm}[1]{\lVert #1 \rVert}
\newcommand{\PiP}{\Pi_P}

\usepackage{hyperref}
\hypersetup{
  colorlinks=true,
  linkcolor=linkcolor,
  citecolor=citecolor,
  urlcolor=linkcolor,
  bookmarksnumbered=true,
  pdfauthor={Lev Petersen},
  pdftitle={Microprice, Bellman Values, and Semi-Markov Reinforcement Learning}
}

\input{rl_note_summary_macros.tex}

\begin{document}

\begin{titlepage}
\centering

\vspace*{1.4cm}

{\Huge\bfseries\color{sectioncolor} Microprice, Bellman Values,\\[0.5em]
Semi-Markov Reinforcement Learning\par}

\vspace{0.45cm}

{\large\color{sectioncolor} A companion note to\\[0.3em]
\textit{Microprice as a Centered Poisson Corrector}\par}

\vspace{0.4cm}
{\color{sectioncolor}\rule{0.42\textwidth}{0.8pt}}

\vspace{1.0cm}

{\Large Lev Petersen\par}

\vspace{1.3cm}

\begin{minipage}{0.86\textwidth}
\setstretch{1.14}
\small
\noindent\textbf{Abstract.}
The starting point is the move-time representation of \cite{Petersen2025}: if the event-time chain splits as $P=Q+T$, then the one-move reward and move-time kernel are
\[
G_1=(I-Q)^{-1}r,
\qquad
B=(I-Q)^{-1}T.
\]
This note shows that the discounted event-time Bellman value is the exact discounted analogue of the same construction. For every $\gamma\in(0,1)$,
\[
V_\gamma=(I-\gamma P)^{-1}r
\]
admits the move-time factorization
\[
V_\gamma=G_{1,\gamma}+B_\gamma V_\gamma,
\qquad
G_{1,\gamma}=(I-\gamma Q)^{-1}r,
\qquad
B_\gamma=\gamma(I-\gamma Q)^{-1}T.
\]
Thus the reinforcement-learning extension is not external to microprice; it is the discounted version of the same operator calculus. The original centered correction is itself a move-time relative value function:
\[
g=G_1-\mu\,\1+Bg,
\qquad
\mu=\pi^T G_1,
\qquad
\pi^T g=0.
\]
So classical microprice may be learned online by differential TD on price-move episodes, while discounted forecasting may be learned by semi-Markov TD on the same episodes. For deterministic horizons,
\[
V^{(H)}=\sum_{k=0}^{H-1}P^k r
\]
is the exact $H$-step Bellman target, and when $P$ is ergodic it decomposes as
\[
V^{(H)}=H\mu_P\,\1+g_P-P^H g_P,
\qquad
g_P=(I-P+\Pi_P)^{-1}(r-\mu_P\1),
\qquad
\mu_P=\pi_P^T r.
\]
The event-time differential value $g_P$ is therefore the exact event-time counterpart of the centered correction in \cite{Petersen2025}.

On the bundled public panels shipped with the workspace (\RLEmpiricalCellCount\ dataset-horizon cells), the fixed-horizon Bellman value beats classical microprice in \RLEmpiricalFiniteWins/\RLEmpiricalCellCount\ cells with geometric-mean grouped-RMSE ratio \RLEmpiricalFiniteGeo, while the horizon-matched discounted value wins in \RLEmpiricalDiscountWins/\RLEmpiricalCellCount\ cells with ratio \RLEmpiricalDiscountGeo. The event-time boundary value $g_P$ is numerically indistinguishable from $G^*$ on the centered Stoikov and WSELOB panels, and improves all six IEX cells with geometric-mean ratio \RLEmpiricalDiffIEXGeo. The semi-Markov formulation also yields a substantial online-learning gain: updating once per price move rather than once per event compresses the update count by a geometric-mean factor \RLEmpiricalMoveCompressionGeo{}x (median \RLEmpiricalMoveCompressionMedian{}x, range \RLEmpiricalMoveCompressionMin{}x--\RLEmpiricalMoveCompressionMax{}x).

\vspace{0.9em}
\noindent\textbf{Keywords:} microprice, reinforcement learning, semi-Markov decision process, temporal-difference learning, value function, successor representation, average reward
\end{minipage}

\vfill

{\small\color{sectioncolor} Draft \textbullet\ \today}

\end{titlepage}

% ============================
\section{Introduction}

The main paper \cite{Petersen2025} identifies classical microprice as a centered Poisson corrector on the move-time chain: after the decomposition $P=Q+T$, the event-time model collapses to a one-move reward $G_1=(I-Q)^{-1}r$ and a move-time kernel $B=(I-Q)^{-1}T$. The undiscounted correction solves
\[
(I-B)g=G_1-\mu\1,
\qquad
\mu=\pi^T G_1,
\qquad
\pi^T g=0,
\]
so the state-dependent part of the long-run predictor is a centered Poisson solution on the post-move chain.

The present note keeps that notation and extends it in two directions that are natural from reinforcement learning. The first is the exact event-time Bellman family on $P$:
\[
V^{(H)}=\sum_{k=0}^{H-1}P^k r,
\qquad
V_\gamma=(I-\gamma P)^{-1}r.
\]
These are the exact fixed-horizon and discounted predictors of future mid-price change. The second is a move-time semi-Markov representation that stays as close as possible to the original paper: the option ``wait until the next mid-price move'' has discounted reward $G_{1,\gamma}=(I-\gamma Q)^{-1}r$ and continuation kernel $B_\gamma=\gamma(I-\gamma Q)^{-1}T$.

The key point is that these are the same value written at different temporal resolutions. The event-time Bellman value $V_\gamma$ can be evaluated on $P$, or equivalently on the move-time semi-Markov chain generated by $B_\gamma$. That equivalence makes the extension seamless with the original microprice paper. It also yields a clean pair of move-time RL updates on the very same state space: differential TD for the original centered microprice correction, and semi-Markov TD for discounted forecasting.

Empirically, the exact fixed-horizon Bellman value remains the strongest forecasting object on the bundled public panels, while the discounted event-time solve remains the cleanest computational object. The event-time differential value $g_P$ sharpens the conceptual bridge back to \cite{Petersen2025}: on the centered Stoikov and WSELOB panels it coincides with $G^*$ to numerical tolerance, whereas on the IEX stress panels it moves in a slightly more favorable direction on all six cells. The move-time updates are also materially more ergonomic: both the differential-TD microprice learner and the discounted SMDP learner update only once per price move, not once per event.

% ============================
\section{Discounted move-time Bellman equations}

Let $X_n$ be the event-time Markov chain of top-of-book states and let $p_n$ be the mid-price. As in \cite{Petersen2025}, write
\[
P=Q+T,
\qquad
r(x):=\E[p_{n+1}-p_n\mid X_n=x].
\]
Here $Q$ collects no-move transitions and $T$ move transitions. Assume $\rho(Q)<1$ so that the fundamental matrix $N=(I-Q)^{-1}$ exists.

\begin{definition}[Classical move-time objects]
The one-move reward and move-time kernel are
\[
G_1:=Nr=(I-Q)^{-1}r,
\qquad
B:=NT=(I-Q)^{-1}T.
\]
These are exactly the objects appearing in \cite{Petersen2025}.
\end{definition}

For $\gamma\in(0,1)$ define the discounted event-time Bellman value
\begin{equation}\label{eq:discounted_event_value}
V_\gamma(x):=\E_x\!\left[\sum_{n=0}^{\infty}\gamma^n(p_{n+1}-p_n)\right].
\end{equation}
The first proposition shows that $V_\gamma$ can be written using the same move-time reduction as microprice.

\begin{proposition}[Discounted move-time factorization]\label{prop:discounted_move_factorization}
For $\gamma\in(0,1)$ define
\[
G_{1,\gamma}:=(I-\gamma Q)^{-1}r,
\qquad
B_\gamma:=\gamma(I-\gamma Q)^{-1}T.
\]
Then $V_\gamma$ satisfies the semi-Markov Bellman equation
\begin{equation}\label{eq:discounted_move_bellman}
V_\gamma = G_{1,\gamma} + B_\gamma V_\gamma.
\end{equation}
Equivalently,
\begin{equation}\label{eq:discounted_move_resolvent}
V_\gamma=(I-B_\gamma)^{-1}G_{1,\gamma}.
\end{equation}
Moreover,
\begin{equation}\label{eq:discounted_move_identity}
I-B_\gamma=(I-\gamma Q)^{-1}(I-\gamma P),
\end{equation}
so \eqref{eq:discounted_move_resolvent} is exactly the same value as
\begin{equation}\label{eq:discounted_event_resolvent}
V_\gamma=(I-\gamma P)^{-1}r.
\end{equation}
\end{proposition}

Equation \eqref{eq:discounted_move_bellman} is the most direct RL bridge from the original paper. The only change from the undiscounted move-time objects $(G_1,B)$ is that every no-move excursion is discounted before it is collapsed.

\begin{remark}[Temporal abstraction]
The option ``wait until the next mid-price move'' is a semi-Markov option in the sense of Sutton, Precup, and Singh \cite{SuttonPrecupSingh1999}. Its option reward is $G_{1,\gamma}$ and its continuation kernel is $B_\gamma$. The original microprice paper corresponds to the undiscounted boundary case $\gamma\uparrow 1$, where $G_{1,\gamma}\to G_1$ and $B_\gamma\to B$ and the constant mode must be centered explicitly.
\end{remark}

\begin{proposition}[Classical microprice as a move-time relative value]\label{prop:move_time_relative_value}
Assume the move-time chain $B$ is ergodic with stationary distribution $\pi$ and set
\[
\mu:=\pi^T G_1,
\qquad
\Pi_B:=\1\pi^T.
\]
Then the centered correction
\begin{equation}\label{eq:move_time_relative_value}
g:=(I-B+\Pi_B)^{-1}(G_1-\mu\1)
\end{equation}
satisfies
\begin{equation}\label{eq:move_time_average_reward_bellman}
g=G_1-\mu\1+Bg,
\qquad
\pi^T g=0.
\end{equation}
So the centered microprice correction of \cite{Petersen2025} is precisely the average-reward relative value function of the move-time chain.
\end{proposition}

This identity is the most seamless RL interpretation of the original paper. The undiscounted object is already a value function; the discounted semi-Markov family simply replaces the average-reward boundary problem by a horizon-aware Bellman objective.

% ============================
\section{Event-time Bellman values and differential limits}

The event-time chain admits an exact finite-horizon Bellman family.

\begin{proposition}[Finite-horizon Bellman values]\label{prop:finite_horizon}
For each integer $H\ge 0$, define
\[
V^{(0)}:=0,
\qquad
V^{(h+1)}:=r+PV^{(h)}.
\]
Then
\begin{equation}\label{eq:finite_horizon_closed}
V^{(H)}=\sum_{k=0}^{H-1}P^k r,
\end{equation}
and, if $R_n:=p_{n+1}-p_n$,
\begin{equation}\label{eq:finite_horizon_expectation}
V^{(H)}(x)=\E_x\!\left[\sum_{n=0}^{H-1}R_n\right].
\end{equation}
Thus $V^{(H)}$ is the exact deterministic-horizon predictor of future mid-price change.
\end{proposition}

When $P$ is ergodic, the same Poisson-corrector structure that appears in \cite{Petersen2025} reappears on event time.

\begin{proposition}[Event-time differential value]\label{prop:event_time_poisson}
Assume $P$ is ergodic with stationary distribution $\pi_P$ and set
\[
\Pi_P:=\1\pi_P^T,
\qquad
\mu_P:=\pi_P^T r.
\]
Define the event-time differential value
\begin{equation}\label{eq:event_time_diff_value}
g_P:=(I-P+\Pi_P)^{-1}(r-\mu_P\1),
\qquad
\pi_P^T g_P=0.
\end{equation}
Then, for every integer $H\ge 0$,
\begin{equation}\label{eq:finite_horizon_decomp}
V^{(H)}=H\mu_P\,\1+g_P-P^H g_P.
\end{equation}
For every $\gamma\in(0,1)$,
\begin{equation}\label{eq:discounted_split_event_time}
V_\gamma
=
\frac{\mu_P}{1-\gamma}\,\1
+
(I-\gamma P)^{-1}(I-\Pi_P)r.
\end{equation}
Hence
\begin{equation}\label{eq:discounted_boundary_event_time}
V_\gamma-\frac{\mu_P}{1-\gamma}\,\1
\longrightarrow g_P
\qquad\text{as }\gamma\uparrow 1.
\end{equation}
\end{proposition}

This is the exact event-time analogue of the main theorem in \cite{Petersen2025}. The move-time correction $G^*$ is the centered long-run value after the $Q/T$ reduction; the event-time correction $g_P$ is the centered long-run value before that reduction.

\begin{remark}[Two boundary objects]
The choice between $G^*$ and $g_P$ is a modelling choice about temporal resolution. If the objective is expressed in moves, $G^*$ is the natural centered boundary object. If the objective is expressed in events or clock-time horizons induced from events, $g_P$ is the natural boundary object. The finite-horizon values $V^{(H)}$ interpolate between short-horizon forecasting and these long-run correctors through \eqref{eq:finite_horizon_decomp}.
\end{remark}

% ============================
\section{Learning rules}

\subsection{Event-time TD and LSTD}

The discounted Bellman equation on event time,
\[
V_\gamma=r+\gamma PV_\gamma,
\]
may be learned online by standard temporal-difference learning \cite{Sutton1988}. With $X_t$ the discretized state and $R_t:=p_{t+1}-p_t$,
\begin{equation}\label{eq:td0}
V_{t+1}(X_t)
\leftarrow
V_t(X_t)+\eta_t\bigl(R_t+\gamma V_t(X_{t+1})-V_t(X_t)\bigr).
\end{equation}
In the tabular case this is a one-pass $O(L)$-memory update. With linear features $\phi(x)$, least-squares TD solves
\begin{equation}\label{eq:lstd}
A_T w_T=b_T,
\qquad
A_T:=\sum_{t<T}\phi_t(\phi_t-\gamma\phi_{t+1})^T,
\qquad
b_T:=\sum_{t<T}\phi_t R_t,
\end{equation}
which is the natural linear policy-evaluation analogue \cite{BradtkeBarto1996}.

The same value also admits the successor-representation factorization \cite{Dayan1993}
\begin{equation}\label{eq:sr}
M_\gamma:=(I-\gamma P)^{-1}=
\sum_{k=0}^{\infty}\gamma^k P^k,
\qquad
V_\gamma=M_\gamma r.
\end{equation}
If rewards are parameterized as $r_\theta=\Phi\theta$, then the successor features are $\Psi_\gamma:=M_\gamma\Phi$ and $V_{\gamma,\theta}=\Psi_\gamma\theta$ \cite{Barreto2018}.

\subsection{Move-time differential TD for classical microprice}

Proposition~\ref{prop:move_time_relative_value} shows that the centered correction of \cite{Petersen2025} is already an average-reward Bellman value on move time. Writing $\Delta P_k:=p_{\tau_{k+1}}-p_{\tau_k}$ for the move increment,
\begin{equation}\label{eq:move_time_relative_expectation}
g(y)=\E\bigl[\Delta P_k-\mu+g(Y_{k+1})\mid Y_k=y\bigr].
\end{equation}
Hence the original microprice correction may be learned online by differential TD. A tabular recursion is
\begin{equation}\label{eq:move_time_differential_td}
\delta_k:=\Delta P_k-\bar\mu_k+h_k(Y_{k+1})-h_k(Y_k),
\end{equation}
\begin{equation}\label{eq:move_time_differential_td_updates}
h_{k+1}(Y_k)\leftarrow h_k(Y_k)+\alpha_k\delta_k,
\qquad
\bar\mu_{k+1}\leftarrow \bar\mu_k+\beta_k\delta_k.
\end{equation}
This is the exact RL version of classical microprice: same move-time chain, same state space, same centered boundary object. In particular, no event-time surrogate is required to learn the predictor online.

\subsection{Move-time semi-Markov TD for discounted forecasting}

For horizon-aware forecasting, replace the average-reward boundary problem by the discounted semi-Markov value. Let $\tau_k$ be the successive mid-price move times, $Y_k:=X_{\tau_k}$ the move-time chain, and $D_k:=\tau_{k+1}-\tau_k$ the move duration. Then the discounted value at move times satisfies the recursion
\begin{equation}\label{eq:smdp_bellman}
W_\gamma(y)=\E\bigl[\Delta P_k+\gamma^{D_k}W_\gamma(Y_{k+1})\mid Y_k=y\bigr].
\end{equation}
The corresponding tabular SMDP-TD(0) update is
\begin{equation}\label{eq:smdp_td}
W_{k+1}(Y_k)
\leftarrow
W_k(Y_k)
+
\eta_k\bigl(\Delta P_k+\gamma^{D_k}W_k(Y_{k+1})-W_k(Y_k)\bigr).
\end{equation}
This update uses the same move-time episodes as differential TD, but it targets the discounted Bellman value rather than the centered long-run corrector. The choice between the two is therefore objective-driven: differential TD learns the classical microprice boundary object, while SMDP-TD learns the horizon-aware discounted predictor.

If one prefers a model-based target, $V^{(H)}$ remains the exact finite-horizon forecast and $V_{\gamma_H}$ with $\gamma_H=1-1/H$ remains the cleanest one-shot surrogate. The move-time learning rules are the online versions of those same objectives.

% ============================
\section{Empirical results}

The empirical role of this note is deliberately narrow. We evaluate the RL extension on the bundled public panels already shipped with the workspace: Stoikov BAC/CVX, IEX BAC/SPY/GLD, and WSELOB PEKAO/PZU/PKOBP/PKNORLEN/KGHM. The grouped conditional-curve metric is the same one used by the existing public scripts.

\begin{table}[t]
\centering
\small
\begin{tabular}{@{}p{4.3cm}ccp{3.2cm}@{}}
\toprule
Predictor & Wins vs. $G^*$ & Geo.-mean RMSE ratio & Interpretation \\
\midrule
Finite-horizon Bellman $V^{(H)}$ & \RLEmpiricalFiniteWins/\RLEmpiricalCellCount & \RLEmpiricalFiniteGeo & exact fixed-horizon target \\
Discounted Bellman $V_{\gamma_H}$ & \RLEmpiricalDiscountWins/\RLEmpiricalCellCount & \RLEmpiricalDiscountGeo & one discounted solve \\
Event-time differential value $g_P$ & IEX 6/6; elsewhere ties & \RLEmpiricalDiffGeo & event-time boundary object \\
\bottomrule
\end{tabular}
\caption{Bundled-data comparison against classical microprice. Ratios below one favor the RL extension. For $g_P$, the non-IEX panels are equal to $G^*$ up to numerical tolerance, so the only visible differences come from the IEX stress cases.}
\label{tab:rl_summary}
\end{table}

Table~\ref{tab:rl_summary} gives the main forecasting message. The exact fixed-horizon Bellman value beats classical microprice in \RLEmpiricalFiniteWins/\RLEmpiricalCellCount\ cells, with geometric-mean RMSE ratio \RLEmpiricalFiniteGeo. The horizon-matched discounted value, with $\gamma_H=1-1/H$, wins in \RLEmpiricalDiscountWins/\RLEmpiricalCellCount\ cells with ratio \RLEmpiricalDiscountGeo. The strongest gains are off the canonical centered regime: PEKAO at ten steps drops from $4.551\times 10^{-3}$ for $G^*$ to $1.003\times 10^{-3}$ for $V^{(H)}$, and PKOBP at ten steps drops from $1.403\times 10^{-3}$ to $2.42\times 10^{-4}$. The clean negative case remains SPY at sixty steps, where $G^*$ has RMSE $8.284\times 10^{-3}$ versus $1.027\times 10^{-2}$ for $V^{(H)}$.

The event-time boundary value $g_P$ does exactly what the theory suggests. On the Stoikov and WSELOB panels, which are already close to the centered regime of \cite{Petersen2025}, $g_P$ and $G^*$ coincide to numerical tolerance. On the six IEX cells, where event time and move time are more visibly separated, $g_P$ improves all six cells with geometric-mean ratio \RLEmpiricalDiffIEXGeo. So $g_P$ is not a replacement for the objective-matched finite-horizon value, but it is the correct event-time boundary object and the cleanest conceptual bridge back to the original paper.

\begin{table}[t]
\centering
\small
\begin{tabular}{lcc}
\toprule
Panel & Geo.-mean events per move & Interpretation \\
\midrule
Stoikov & \RLEmpiricalMoveCompressionStoikovGeo{}x & modest temporal aggregation \\
IEX & \RLEmpiricalMoveCompressionIEXGeo{}x & moderate update compression \\
WSELOB & \RLEmpiricalMoveCompressionWSEGeo{}x & very sparse move times \\
Overall & \RLEmpiricalMoveCompressionGeo{}x & median \RLEmpiricalMoveCompressionMedian{}x \\
\bottomrule
\end{tabular}
\caption{Online-learning compression from move-time TD updates. One move-time update --- differential TD for classical microprice or SMDP-TD for discounted forecasting --- replaces one event-time TD step for each no-move event absorbed into the current price-move episode.}
\label{tab:compression}
\end{table}

Table~\ref{tab:compression} records the ergonomic gain behind the move-time RL formulation. Both the differential-TD microprice learner and the discounted SMDP learner update only once per price move, so the update count is compressed by a geometric-mean factor \RLEmpiricalMoveCompressionGeo{}x overall, with median \RLEmpiricalMoveCompressionMedian{}x and range \RLEmpiricalMoveCompressionMin{}x to \RLEmpiricalMoveCompressionMax{}x. This matters most on the sparse-move panels: the WSELOB names sit at \RLEmpiricalMoveCompressionWSEGeo{}x on average, and PEKAO reaches \RLEmpiricalMoveCompressionMax{}x. In other words, the move-time RL rules are not just notational reformulations. They are substantial variance-reduction and throughput improvements whenever no-move events dominate the tape.

The discounted linear solve remains the cleanest model-based object as well. On the bundled models, solving $(I-\gamma P)V=r$ is \RLEmpiricalDiscountSpeedMean{}x cheaper on average than the classical microprice solve stage. So the practical message is split cleanly: $V^{(H)}$ is the exact forecasting target, $V_{\gamma_H}$ is the simplest fast surrogate, and the move-time SMDP formulation is the most ergonomic online learner.

% ============================
\section{Conclusion}

The reinforcement-learning extension of microprice is best understood as an extension of the original move-time operator calculus, not as a separate theory. The same decomposition $P=Q+T$ that produces the undiscounted pair $(G_1,B)$ also produces the discounted pair $(G_{1,\gamma},B_\gamma)$, and the exact discounted Bellman value may be written either on event time or on the corresponding semi-Markov move-time chain.

This viewpoint keeps the mathematics and the implementation aligned. The exact fixed-horizon Bellman value $V^{(H)}$ matches the forecasting objective, the discounted value $V_\gamma$ gives the cleanest one-shot solve, the event-time differential value $g_P$ is the correct event-time boundary object, the move-time differential recursion learns the classical microprice correction itself, and the move-time SMDP recursion learns the horizon-aware discounted predictor on the same state space used in \cite{Petersen2025}. The empirical picture is correspondingly simple: fixed-horizon Bellman values produce the largest forecasting gains, while move-time RL updates deliver the largest ergonomic gain.

\appendix

% ============================
\section{Proofs}

\begin{proof}[Proof of Proposition~\ref{prop:discounted_move_factorization}]
Partition the discounted return at the first price move. If the chain makes $n$ no-move transitions before the move step, then the contribution of the first episode is $\gamma^n Q^n r$, and the post-move continuation to state $y$ contributes $\gamma^{n+1}(Q^nT)_{xy}V_\gamma(y)$. Summing over $n\ge 0$ gives
\[
V_\gamma
=
\sum_{n\ge 0}\gamma^n Q^n r
+
\sum_{n\ge 0}\gamma^{n+1}Q^n T V_\gamma
=
(I-\gamma Q)^{-1}r+\gamma(I-\gamma Q)^{-1}T V_\gamma,
\]
which is \eqref{eq:discounted_move_bellman}. Rearranging yields \eqref{eq:discounted_move_resolvent}. Finally,
\[
(I-\gamma Q)(I-B_\gamma)
=
I-\gamma Q-\gamma T
=
I-\gamma P,
\]
which is \eqref{eq:discounted_move_identity}. Multiplying \eqref{eq:discounted_move_resolvent} by $I-\gamma P$ therefore gives \eqref{eq:discounted_event_resolvent}.
\end{proof}

\begin{proof}[Proof of Proposition~\ref{prop:move_time_relative_value}]
By definition, $g$ solves
\[
(I-B+\Pi_B)g=G_1-\mu\1.
\]
Because $\Pi_B g=\1\pi^T g$, the normalization $\pi^T g=0$ reduces this to
\[
(I-B)g=G_1-\mu\1,
\]
which rearranges to \eqref{eq:move_time_average_reward_bellman}. Conversely, if \eqref{eq:move_time_average_reward_bellman} holds together with $\pi^T g=0$, then applying $\pi^T$ gives $\mu=\pi^T G_1$, and adding $\Pi_B g=0$ recovers \eqref{eq:move_time_relative_value}.
\end{proof}

\begin{proof}[Proof of Proposition~\ref{prop:finite_horizon}]
The recursion implies inductively that
\[
V^{(H)}=r+Pr+\cdots+P^{H-1}r=\sum_{k=0}^{H-1}P^k r,
\]
which is \eqref{eq:finite_horizon_closed}. Since $P^k r(x)=\E_x[R_k]$, summing over $k=0,\dots,H-1$ gives \eqref{eq:finite_horizon_expectation}.
\end{proof}

\begin{proof}[Proof of Proposition~\ref{prop:event_time_poisson}]
Since $g_P$ solves $(I-P)g_P=r-\mu_P\1$, we have $r=\mu_P\1+g_P-Pg_P$. Therefore
\[
V^{(H)}
=
\sum_{k=0}^{H-1}P^k r
=
\sum_{k=0}^{H-1}P^k(\mu_P\1+g_P-Pg_P)
=
H\mu_P\1+g_P-P^H g_P,
\]
which is \eqref{eq:finite_horizon_decomp}. For the discounted identity, decompose
\[
r=\Pi_P r+(I-\Pi_P)r=\mu_P\1+(I-\Pi_P)r.
\]
Because $P\1=\1$ and $P\Pi_P=\Pi_P$,
\[
(I-\gamma P)^{-1}\Pi_P r
=
\sum_{k=0}^{\infty}\gamma^k\Pi_P r
=
\frac{\mu_P}{1-\gamma}\,\1.
\]
This gives \eqref{eq:discounted_split_event_time}. On the centered subspace $(I-\Pi_P)\R^L$, the powers of $P$ converge to zero, so the Neumann series converges to $(I-P+\Pi_P)^{-1}(I-\Pi_P)r=g_P$, which yields \eqref{eq:discounted_boundary_event_time}.
\end{proof}

\newpage
\begingroup
\small
\setlength{\itemsep}{3pt plus 1pt minus 1pt}
\setlength{\parsep}{0pt}
\begin{thebibliography}{99}

\bibitem{Petersen2025}
L.~Petersen, ``Microprice as a Centered Poisson Corrector,'' preprint, 2025.

\bibitem{Stoikov2018}
S.~Stoikov, ``The Micro-Price: A High Frequency Estimator of Future Prices,'' \textit{Quantitative Finance}, vol.~18, no.~12, pp.~1959--1966, 2018. \href{https://doi.org/10.1080/14697688.2018.1489139}{doi:10.1080/14697688.2018.1489139}.

\bibitem{Sutton1988}
R.~S.~Sutton, ``Learning to Predict by the Methods of Temporal Differences,'' \textit{Machine Learning}, vol.~3, no.~1, pp.~9--44, 1988. \href{https://doi.org/10.1023/A:1022633531479}{doi:10.1023/A:1022633531479}.

\bibitem{BradtkeBarto1996}
S.~J.~Bradtke and A.~G.~Barto, ``Linear Least-Squares Algorithms for Temporal Difference Learning,'' \textit{Machine Learning}, vol.~22, pp.~33--57, 1996. \href{https://doi.org/10.1023/A:1018056104778}{doi:10.1023/A:1018056104778}.

\bibitem{Dayan1993}
P.~Dayan, ``Improving Generalization for Temporal Difference Learning: The Successor Representation,'' \textit{Neural Computation}, vol.~5, no.~4, pp.~613--624, 1993. \href{https://doi.org/10.1162/neco.1993.5.4.613}{doi:10.1162/neco.1993.5.4.613}.

\bibitem{SuttonPrecupSingh1999}
R.~S.~Sutton, D.~Precup, and S.~Singh, ``Between MDPs and Semi-MDPs: A Framework for Temporal Abstraction in Reinforcement Learning,'' \textit{Artificial Intelligence}, vol.~112, no.~1--2, pp.~181--211, 1999. \href{https://doi.org/10.1016/S0004-3702(99)00052-1}{doi:10.1016/S0004-3702(99)00052-1}.

\bibitem{TsitsiklisVanRoy1999}
J.~N.~Tsitsiklis and B.~Van Roy, ``Average Cost Temporal-Difference Learning,'' \textit{Automatica}, vol.~35, no.~11, pp.~1799--1808, 1999. \href{https://doi.org/10.1016/S0005-1098(99)00115-1}{doi:10.1016/S0005-1098(99)00115-1}.

\bibitem{TsitsiklisVanRoy2002}
J.~N.~Tsitsiklis and B.~Van Roy, ``On Average versus Discounted Reward Temporal-Difference Learning,'' \textit{Machine Learning}, vol.~49, pp.~179--191, 2002. \href{https://doi.org/10.1023/A:1017994911517}{doi:10.1023/A:1017994911517}.

\bibitem{Barreto2018}
A.~Barreto, D.~Borsa, J.~Quan, T.~Schaul, D.~Silver, M.~Hessel, D.~Mankowitz, A.~Zidek, and R.~Munos, ``Transfer in Deep Reinforcement Learning Using Successor Features and Generalised Policy Improvement,'' in \textit{Proceedings of the 35th International Conference on Machine Learning}, PMLR~80, pp.~501--510, 2018.

\end{thebibliography}
\endgroup

\end{document}
